---
title: "p8105_hw6_dl3757"
author: "Dang Lin dl3757"
date: "2024-12-02"
output: github_document
---

```{r message = FALSE, warning = FALSE}
# Import the libraries
library(tidyverse)
library(purrr)
library(broom)
library(modelr)
```

# Problem 1
```{r message = FALSE, warning = FALSE}
# Load and clean the dataset
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r message = FALSE, warning = FALSE}
# Set seed for reproducibility
set.seed(1)

# Bootstrap
boot_straps <- 
  weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin, data = df)),
    results = map(models, broom::tidy), 
    r_squared = map(models, \(model) broom::glance(model) %>% pull(r.squared)),
    log_beta_product = map(models, \(model) {
      coefs <- broom::tidy(model)
      beta_0 <- coefs %>% filter(term == "(Intercept)") %>% pull(estimate)
      beta_1 <- coefs %>% filter(term == "tmin") %>% pull(estimate)
      log(beta_0 * beta_1)
      })
    ) %>% 
  select(-strap, -models)
```

```{r message = FALSE, warning = FALSE}
# Density plot of R-squared
boot_straps %>%
  ggplot(aes(x = r_squared %>% unlist())) +
  geom_density() +
  labs(title = "Density Plot of R-squared", 
       x = "R-squared", 
       y = "Density") + 
  theme_minimal()
```

The most likely R-squared value is approximately 0.913, indicating a high proportion of variation explained by the model. The plot shows that the distribution is unimodal and approximately symmetric, with R-squared values ranging from 0.88 to 0.94. 

```{r message = FALSE, warning = FALSE}
# Density plot of log(beta0 * beta1)
boot_straps %>%
  ggplot(aes(x = log_beta_product %>% unlist())) +
  geom_density() + 
  labs(title = "Density Plot of log(beta0 * beta1)",
    x = "log(beta0 * beta1)",
    y = "Density") + 
  theme_minimal()
```

The most likely value of log(beta0 * beta1) is approximately 2.00, indicating the central tendency of the distribution. The plot shows that the distribution is unimodal and approximately symmetric, with log(beta0 * beta1) values ranging from 1.94 to 2.10.

```{r message = FALSE, warning = FALSE}
# 95% Confidence Interval for R-squared
boot_straps %>%
  pull(r_squared) %>% 
  unlist() %>%
  quantile(c(0.025, 0.975), na.rm = TRUE) %>% 
  knitr::kable(col.names = "95% Confidence Interval for R-squared", 
               digits = 4)
```

The 95% Confidence Interval for R-squared is (0.8937, 0.9271).

```{r message = FALSE, warning = FALSE}
# 95% Confidence Interval for log(beta0 * beta1)
boot_straps %>%
  pull(log_beta_product) %>% 
  unlist() %>%
  quantile(c(0.025, 0.975), na.rm = TRUE) %>% 
  knitr::kable(col.names = "95% Confidence Interval for log(beta0 * beta1)", 
               digits = 4)
```

The 95% Confidence Interval for log(beta0 * beta1) is (1.9649, 2.0589).
